---
title: ""
output:
  pdf_document: default
  html_document: default
date: ""
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Code Blocks
## Howe and Nelson *in prep.*
### Please cite this manuscript if you use our generic code examples.


**Code Block 1:** Generic example of model construction using **A)** *S3* objects (mlr) and **B)** *R6* objects (mlr3). In both approaches we are randomly drawing a training (train.set) and test (test.set) dataset from our full dataset (full.dataset). The model stacking ensemble that we use in **B** requires *mlr3pipelines*.

**1.A**

```{r 1.A, eval=FALSE, attr.source='.numberLines'}

#### S3: mlr  ####

learner.randomforest = setHyperPars(
              makeLearner("classif.ranger", 
              num.threads=12, 
              predict.type = "prob", 
              importance="impurity", 
              predict.threshold = 0.5), 
              par.vals = list(num.trees=250, verbose=TRUE))

learner.XGBoost = setHyperPars(
              makeLearner("classif.xgboost", 
              nthread=12, 
              predict.type = "prob", 
              predict.threshold = 0.5), 
              par.vals = list(nrounds = 250,
              max_depth = 8,
              alpha=0.022,
              eta = 0.36, 
              lambda = 0.1))

ensemble = makeStackedLearner(base.learners = list(
    learner.randomforest, 
    learner.XGBoost),
    method="average",
    predict.type = "prob")

train.set <- full.dataset %>%
    group_by(mpb) %>%
    slice_sample(n=50000) %>%
    mutate(mpb=as.factor(mpb)) %>%
    as.data.frame()
  
test.set <- full.dataset %>%
    group_by(mpb) %>%
    slice_sample(n=50000) %>%
    mutate(mpb=as.factor(mpb)) %>%
    as.data.frame()
  
model.train <- train.set %>%
    makeClassifTask(data=., target="mpb", positive="1") %>%
    train(ensemble, .)

model.predict <- test.set %>%
    makeClassifTask(data=., target="mpb", positive="1") %>%
    predict(model.train, .)
```

**1.B**

```{r 1.B, eval=FALSE, attr.source='.numberLines'}

#### R6: mlr3 ####

learner.randomforest <- lrn(
  "classif.ranger", 
  num.threads=12, 
  num.trees = 250, 
  importance="impurity",
  predict_type="prob")

learner.XGBoost <- lrn(
  "classif.xgboost",
  nthread=12,
  alpha=0.022,
  eta = 0.36, 
  lambda = 0.1,
  max_depth=8,
  nrounds=250,
  predict_type="prob")

train.set <- full.dataset %>%
    group_by(mpb) %>%
    slice_sample(n=50000) %>%
    as.data.frame()
  
test.set <- full.dataset %>%
    group_by(mpb) %>%
    slice_sample(n=50000) %>%
    as.data.frame()

training.task <- as_task_classif(x=train.set, target="mpb")  
testing.task <-  as_task_classif(x=test.set, target="mpb")  

ensemble.pipeline <- 
  gunion(list(
    PipeOpLearnerCV$new(learner.randomforest, id = "xgb"),
    PipeOpLearnerCV$new(learner.XGBoost, id = "rf")
  ))

ensemble.average <- LearnerClassifAvg$new(id = "classif")

ensemble.model <- ensemble.pipeline %>>%
  PipeOpFeatureUnion$new(2, id = "avg") %>>%
  ensemble.average

ensemble <- GraphLearner$new(ensemble.model, predict_type="prob")
model.train <- ensemble$train(training.task)
model.test <- ensemble$predict(test.task)

```

**Code Block 2:** Generic example of hyperparameter tuning using the default tuning space for XGBoost. We constrained the number of rounds to 250-500 and the max depth to 5-10 based on processing limitations.

```{r 2, eval=FALSE, attr.source='.numberLines'}

#### tuning space ###

xgboost.lts <- lts("classif.xgboost.default",
              nrounds=to_tune(250, 500),
              max_depth=to_tune(5,10),
              subsample=NULL,
              colsample_bylevel=NULL,
              colsample_bytree=NULL)

#### tuner ####

xgboost.ti <- ti(
  task=training.task,
  learner=lrn("classif.xgboost", nthread=12),
  resampling=rsmp("cv", folds=3),
  measures = msr("classif.ce"),
  terminator=trm("evals", n_evals=10),
  search_space = c.xgb.lts
)

#### grid search  ####

xgboost.tnr <- tnr("grid_search", batch_size=50)

#### optimize ####

xgboost.tnr$optimize(xgboost.ti)

```

**Code Block 3:** Generic example of forward selection using **A)** mlr3fselect; and **B)** our forward selection algorithm. Our variable selection function 'var.selex' calls the function 'ensemble.model' which simply takes a list of training and testing datasets (one for each predictor), fits a model based on the training dataset and calculates perforamnce based on the testing dataset. 

**3.A**

```{r, eval=FALSE, attr.source='.numberLines'}

model.fselect = fselect(
  fselector = fs("sequential", max_features=10),
  task = training.task,
  learner = ensemble,
  resampling = rsmp("holdout"),
  measure = msr("classif.ce"))

```

**3.B**

```{r, eval=FALSE, attr.source='.numberLines'}

ensemble.model <- function(df.train, df.test){
  
  df.prdctr <- df.train %>% select(predictor) %>% distinct() %>% pull()
  df.trn <- df.train %>% select(-predictor)
  df.tst <- df.test %>% select(-predictor)
  
  sample.tsk <- as_task_classif(x=df.trn, target="mpb")
  sample.tst <- as_task_classif(x=df.tst, target="mpb")

  m.e.cv = GraphLearner$new(m.1, predict_type="prob")  
  m.train <- m.e.cv$train(sample.tsk)
  m.prd <- m.e.cv$predict(sample.tst)
  
  m.prf <- data.frame(
    acc = m.prd$score(msr("classif.acc")),
    auc = m.prd$score(msr("classif.auc")),
    ce = m.prd$score(msr("classif.ce")),
    fdr = m.prd$score(msr("classif.fdr")),
    fnr = m.prd$score(msr("classif.fnr")),
    fomr = m.prd$score(msr("classif.fomr")),
    fpr = m.prd$score(msr("classif.fpr")),
    tnr = m.prd$score(msr("classif.tnr")),
    tpr = m.prd$score(msr("classif.tpr"))) %>%
    as_tibble() %>% bind_cols(t(m.train$timings) %>% as_tibble()) %>% 
    mutate(predictor=df.prdctr)

  return(m.prf)
  
}

`%ni%` <- Negate(`%in%`)

var.selex <- function(train.df, test.df, max.feat=14, cutoff=0.8, iter=j){
  
  #### correlation matrix ####
  
  corr.mat <- bind_rows(
    train.df %>% 
    bind_rows(test.df) %>%
    select(-mpb) %>% 
    as.matrix() %>% 
    cor(., method="pearson") %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    pivot_longer(-rowname) %>%
    mutate(method="pearson"),
    train.df %>% 
    bind_rows(test.df) %>%
    select(-mpb) %>% 
    as.matrix() %>% 
    cor(., method="spearman") %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    pivot_longer(-rowname) %>%
    mutate(method="spearman")) %>%
    mutate(abs.value=abs(value)) %>%
    select(var.1=rowname,
           var.2=name,
           method, abs.value) %>%
    pivot_wider(values_from=abs.value, names_from=method)
  
  #### result.grid  ####
  
  result.grid <- list()

  for(i in 1:max.feat){
    
  print(paste0("#### model ", i, " ####"))  
    
  if(i == 1){  

  train.split <- train.df %>%
  pivot_longer(-mpb, names_to="predictor", values_to="obs") %>%
  split(.$predictor, drop=TRUE) 
  
  test.split <- test.df %>%
  pivot_longer(-mpb, names_to="predictor", values_to="obs") %>%
  split(.$predictor)    
  
  model <- map2(.x=train.split, .y=test.split, .f=ensemble.model) %>% bind_rows()  
  
result.grid[[i]] <- p.s <- model %>% mutate(n.feat=i)
   
  }else{
    
  p.s <- result.grid %>% bind_rows() %>% group_by(n.feat) %>%
    slice_min(order_by=ce) %>% ungroup()
  
  crm <- corr.mat %>%
    filter(var.1 %in% p.s$predictor) %>%
    filter(pearson >= cutoff | spearman >= cutoff)
  
  can.var <- corr.mat %>% filter(var.2 %ni% crm$var.2) %>%
    select(var.2) %>% distinct()
    
  if(nrow(can.var) > 0){
  
  train.split <- train.df %>%
    select(mpb, contains(p.s$predictor), contains(can.var$var.2)) %>%
    pivot_longer(!c(mpb, contains(p.s$predictor)), names_to="predictor", values_to="obs") %>%
    split(.$predictor, drop=TRUE) 
  
  test.split <- test.df %>%
    select(mpb, contains(p.s$predictor), contains(can.var$var.2)) %>%
    pivot_longer(!c(mpb, contains(p.s$predictor)), names_to="predictor", values_to="obs") %>%
    split(.$predictor, drop=TRUE) 
  
  model <- map2(.x=train.split, .y=test.split, .f=ensemble.model) %>% bind_rows()
  
  result.grid[[i]] <- p.s <- model %>% mutate(n.feat=i)
  }
  
  print("done")
  
  }  
    
    
  }
  
  results <- result.grid %>% bind_rows() %>%
    write_csv(paste0("var.selex/vs.", iter, ".csv"))
  
}

```

**Code Block 4:** Methods and code for model interpretation.

```{r, eval=FALSE, attr.source='.numberLines'}

#### explain_mlr3 allows us to easily interpret mlr3 based models

mpb.exp <- explain_mlr3(model=ensemble,
                        data=test.set %>% select(-mpb),
                           y=test.set %>% select(mpb)) 

#### model_profile allows us to create accumulated local dependence profiles; 'variable_splits_with_obs=TRUE' is an additional useful option as it predicts all existing unique values in the test data set (do not use if values are not rounded).

mpb.mp <- model_profile(mpb.exp, type="accumulated", N=1000, variable_splits_with_obs=TRUE)

#### relativizing profiles by mean predictions (relative mean predictions)
mean.exp <- mean(mpb.exp$y_hat)

x.mt <- mpb.mp$agr_profiles %>%
    janitor::clean_names() %>%
    as_tibble() %>%
    mutate(y0=mean.exp) %>%
    select(variable=vname, x=x, yhat=yhat, y0) %>%
    mutate(yt=yhat-y0, iter=iter) 

#### model performance based on:

#### the model explainer
model.performance <- model_performance(mpb.exp)
  
#### directly from mlr3
model.performance <- data.frame(
    acc = model.test$score(msr("classif.acc")),
    auc = model.test$score(msr("classif.auc")),
    ce = model.test$score(msr("classif.ce")),
    fdr = model.test$score(msr("classif.fdr")),
    fnr = model.test$score(msr("classif.fnr")),
    fomr = model.test$score(msr("classif.fomr")),
    fpr = model.test$score(msr("classif.fpr")),
    tnr = model.test$score(msr("classif.tnr")),
    tpr = model.test$score(msr("classif.tpr"))) %>%
  mutate(iter=iter)

#### variable importance

variable.importance <- bind_rows(
    model_parts(mpb.exp) %>% as.data.frame())
  
#### interactions (h2)
friedmans.h2 <- hstats::hstats(mpb.exp)
summary.h2 <- summary(friedmans.h2)

two.way <- summary.h2$h2_pairwise$M %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    select(inx=rowname, h2=V1) %>%
    mutate(type="two-way")
    
one.way <- summary.h2$h2_overall$M %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    select(inx=rowname, h2=V1) %>%
    mutate(type="one-way") 
  
overall <- data.frame(inx="overall", h2=summary.h2$h2$M, type="overall")

```





